Comparative analysis of PMV Models accuracy implemented in the ISO 7730:2005 and ASHRAE 55:2023

### 2025-01-20 Manuscript/src/sections/methodology.tex
review: fixed typos

Both the \gls{55} and \gls{7730} specify a set of applicability limits.
We hence filtered out those data points that did not meet the inclusion criteria of both standards.
The rationale is that the models' accuracy should only be tested within their applicability ranges.
The limits we used are as follows:
\num{10}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius},
\num{10}~$\leq$~\ac{tr}~$\leq$~\qty{40}{\celsius},
\num{0}~$\leq$~\ac{vr}~$\leq$~\qty{1}{\m\per\s},
\num{0}~$\leq$~\ac{clo}~$\leq$~\qty{1.5}{clo},
\num{0}~$\leq$~water vapor partial pressure~$\leq$~\qty{2000}{\pascal},
and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met}.
These criteria removed an additional \var{entries_db_filtered_by_limit_inputs} datapoints.
Consequently, the final dataset used for the analysis contains \var{entries_db_used} entries.
The adjusted total clothing insulation and relative airspeed, as required by both standards, were used to calculate the \ac{pmv} and \ac{pmv-ce} values.

@@

In addition, it should also be noted that while all the data included in the \ac{db2} are published in peer-reviewed papers, not all entries were collected using the same standardised methodology.
For example, while both thermal comfort standards recommend measuring \ac{v} at three different heights, this was not always done.
The absence of standardized measurement protocols, the use of inaccurate or un-calibrated sensors, and the use of different reference tables to estimate \ac{clo} and \ac{met} may all introduce an error in the measurements and, consequently, in the results of both models.
This issue has been already evidenced in the literature by \mycite{dambrosio_alfano_role_2011}.
They showed that the reliability of the \ac{pmv} model is strongly affected by the accuracy of the input data.
In our analysis, we compensate this issue by:
\begin{enumerate}[ {}1{)} ]
    \item comparing the results of the two models using the same dataset; hence, the different in the output is mainly due to the different formulations of the models;
    \item using a large dataset to reduce the bias introduced by the lack of adherence to the measurement protocol recommended by both standards since it can be assumed that measurement errors random across different studies.
    \item using only datasets that have been published in peer-reviewed journals.
%    \item excluding all those studies that did not measure all the environmental parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
\end{enumerate}

@@

To remove this rounding error, we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
However, on their own, are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002} since the model is not expected to predict the exact \ac{tsv} of each participant.
These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{vr}).
If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
The standard deviation would reflect the combined effect of the people's individual differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
This is a similar assumption to the one used by the \ac{pmv} model which ignores individual differences and calculates the average thermal sensation of a `typical' average person.
According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{\pm 0.25}.
However, we believe that this range is too narrow since \ac{tsv} is commonly measured using an ordinal scale of integers and even when environmental parameters are measured within the accuracy of the standards, the output of the model may vary by more than \num{\pm 0.25}~\cite{dambrosio_alfano_role_2011}.
Consequently, we expanded the range to \num{\pm 0.5}.
This is needed to ensure that each \ac{pmv} value can be matched to a \ac{tsv}.
We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range.
Ideally, we would want to have the \ac{pmv} to have higher precision.
however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale and by the accuracy of the inputs.
When the aim is to predict the average thermal sensation of a large group of occupants, binning data by the independent variables is a better representation.
Grouping the participants' responses by their \ac{tsv} mixes people who reported the same \ac{tsv} vote but could have been in different locations and moments of the year and day.

>>>>

Both the \gls{55} and \gls{7730} specify a set of applicability limits.
We hence filtered out those data points that did not meet the inclusion criteria of both standards.
The rationale is that the models' accuracy should only be tested within their applicability ranges.
The limits we used are as follows:
\num{10}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius},
\num{10}~$\leq$~\ac{tr}~$\leq$~\qty{40}{\celsius},
\num{0}~$\leq$~\ac{vr}~$\leq$~\qty{1}{\m\per\s},
\num{0}~$\leq$~\ac{clo}~$\leq$~\qty{1.5}{clo},
\num{0}~$\leq$~water vapor partial pressure~$\leq$~\qty{2000}{\pascal},
and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met}.
These criteria removed an additional \var{entries_db_filtered_by_limit_inputs} data points.
Consequently, the final dataset used for the analysis contains \var{entries_db_used} entries.
The adjusted total clothing insulation and relative airspeed, as required by both standards, were used to calculate the \ac{pmv} and \ac{pmv-ce} values.

@@

In addition, it should also be noted that while all the data included in the \ac{db2} are published in peer-reviewed papers, not all entries were collected using the same standardized methodology.
For example, while both thermal comfort standards recommend measuring \ac{v} at three different heights, this was not always done.
The absence of standardized measurement protocols, the use of inaccurate or uncalibrated sensors, and the use of different reference tables to estimate \ac{clo} and \ac{met} may all introduce an error in the measurements and, consequently, in the results of both models.
This issue has already been evidenced in the literature by \mycite{dambrosio_alfano_role_2011}.
They showed that the reliability of the \ac{pmv} model is strongly affected by the accuracy of the input data.
In our analysis, we compensate this issue by:
\begin{enumerate}[ {}1{)} ]
    \item comparing the results of the two models using the same dataset; hence, the difference in the output is mainly due to the different formulations of the models;
    \item using a large dataset to reduce the bias introduced by the lack of adherence to the measurement protocol recommended by both standards, since it can be assumed that measurement errors are random across different studies.
    \item using only datasets that have been published in peer-reviewed journals.
%    \item excluding all those studies that did not measure all the environmental parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
\end{enumerate}

@@

To remove this rounding error, we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
However, on their own, are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002} since the model is not expected to predict the exact \ac{tsv} of each participant.
These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{vr}).
If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
The standard deviation would reflect the combined effect of the people's differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
This is a similar assumption to the one used by the \ac{pmv} model, which ignores individual differences and calculates the average thermal sensation of a `typical' average person.
According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{\pm 0.25}.
However, we believe that this range is too narrow since \ac{tsv} is commonly measured using an ordinal scale of integers and even when environmental parameters are measured within the accuracy of the standards, the output of the model may vary by more than \num{\pm 0.25}~\cite{dambrosio_alfano_role_2011}.
Consequently, we expanded the range to \num{\pm 0.5}.
This is needed to ensure that each \ac{pmv} value can be matched to a \ac{tsv}.
We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range.
Ideally, we would want to have the \ac{pmv} to have higher precision.
However, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale and by the accuracy of the inputs.
When the aim is to predict the average thermal sensation of a large group of occupants, binning data by the independent variables is a better representation.
Grouping the participants' responses by their \ac{tsv} mixes people who reported the same \ac{tsv} vote but could have been in different locations and moments of the year and day.

### 2025-01-20 Manuscript/src/sections/results.tex
review: fixed typos

Figure~\ref{fig:dist_other_data} shows the distribution of age, height, weight, and running mean outdoor temperature.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/dist_other_data}
    \caption{Distribution of age, height, weight, and running mean outdoor temperature.}
    \label{fig:dist_other_data}
\end{figure*}
Less than half (\var{entries_with_sex_info}) of the total entries have information about the participant's sex, and are almost equally distributed among males (\qty{52}{\percent}) and females.
Ages are not normally distributed, the same is true for the running mean outdoor temperature.
About half of the participants (\var{percentage_age_20_35}) were aged between \num{20} and \num{35} years old, and only \var{percentage_age_over_60} were older than 60.
This limits our analysis to young adults and does not allow us to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} models in predicting thermal sensation for children or older adults.
The \ac{db2} does not contain information about the health status of the participants.
Approximately \var{running_mean_10_25} of the running mean outdoor temperature values are between \qtyrange{10}{25}{\celsius}.

@@

    Each bin in the right figure has more data points than in the left figure, since less data on \ac{tpv} were available.}

@@

Fanger, in his seminal book on thermal comfort wrote that ``The dissatisfied are defined here as those who vote \num{-2} (cool) or \num{-3} (cold), \num{2} (warm) or \num{3} (hot).

@@

\subsection{Model Bias}\label{subsec:model-bias}
The intended aim of the \ac{pmv} model is not to accurately predict each individual thermal response from participants.
The \ac{pmv} model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.
The results presented so far are primarily comparing individual \ac{tsv} to \ac{pmv}.
This type of analysis does not conclusively disprove the inefficacy of the \ac{pmv} models in predicting the average thermal sensation of a large group of occupants.
Moreover, grouping the results into discrete categories introduces rounding errors since some participants reported \ac{tsv} on a continuous scale and the \ac{pmv} value is a continuous variable.
To compensate for this we plotted the \ac{pmv} and \ac{pmv-ce} values as a function of \ac{tsv} in Figure~\ref{fig:bubble_models_vs_tsv} and we plotted a \ac{lowess} curve.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/bubble_models_vs_tsv}
    \caption{The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data.
    Raw data were then binned and a bubble chart (circle area is proportional to the number of votes in that bin) was superimposed over the regression curve to aid the visualization of a large dataset.
    The brown dashed line represents the identity line where the slope is 1 and the intercept is 0.}
    \label{fig:bubble_models_vs_tsv}
\end{figure*}
The curve, in the Figure, is calculated using the individual data and not the binned data.
We only binned the data to aid the visualization of a large dataset.
For the model to accurately predict people’s thermal sensation, the regression line (black continuous line) should pass through the origin of the Cartesian plane and have a slope of 1.
This ideal prediction is referred to as the identity line, shown as the brown dashed line in the figure.
The intercepts for the \ac{pmv} is \var{intercept_all_pmv} and for the \ac{pmv-ce} is \var{intercept_all_pmv_ce} which means that the model has a bias of less than one-quarter of a thermal sensation interval.
% todo check data manually
This result is aligned with those presented in Figure~\ref{fig:bar_stacked_model_accuracy} where the \ac{pmv} and \ac{pmv-ce} models can predict people who reported to be neutral \qty{51}{\percent} and \qty{58}{\percent} of the times, respectively.
Both \ac{pmv} formulations always under-predict the thermal sensation of people who reported to be on the extremes of the \ac{tsv};
the slope of the curve is significantly lower than 1.

\subsubsection{Model Overall Bias}\label{subsubsec:model-overall-bias}
We calculated the overall bias of the model as previously done by \mycite{Humphreys2002} by subtracting the \ac{pmv} and \ac{pmv-ce} models prediction from the self-reported \ac{tsv}.
Figure~\ref{fig:hist_discrepancies} contains the summary statistics.
% todo check data manually
Overall both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/hist_discrepancies}
    \caption{Overall bias of the \ac{pmv} and \ac{pmv-ce} models for the subset of data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
    The x-axis shows the discrepancy (bias) between the predicted and self-reported \ac{tsv}.
    We have binned the data into intervals of equal width.
    We colored the bars in the range from \num{-.5} to \num{.5} to highlight the number of responses that are within this range.
    We report the median and inter-quantile ranges since data are not normally distributed.
    }
    \label{fig:hist_discrepancies}
\end{figure*}
When including only the data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} the median bias of the \ac{pmv} model reduces to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improves.
The median bias for the \ac{pmv-ce} also decreases to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0} but it is higher than the one for the \ac{pmv}.

\subsubsection{Model Bias as Function of Five Input Variables}\label{subsubsec:model-bias-variable}
Here we present the model bias as a function of \ac{tdb}, \ac{vr}, \ac{rh}, and \ac{met}.
Results are shown in Figure~\ref{fig:bias_models}.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/bias_models}
    \caption{Bias of the \ac{pmv} and \ac{pmv-ce} models plotted as a function of the input variables.
    The x ticks labels show the bin range.
    The square bracket($]$) means that the right value is included in the range.
    The red dashed lines show the acceptable range of bias.
    Each bin has a different number of data points.
    }
    \label{fig:bias_models}
\end{figure*}
In each sub-figure, we have binned the results into discrete intervals.
The x-axis tick label shows the bin range.
All the boxen-plots have the same area to aid visualization, even though the number of points in each bin is not constant.
The results depict that the use of the \ac{pmv-ce} model does not improve the bias of the \ac{pmv} results across all input variables.
% todo check data manually
On the contrary, the bias of the \ac{pmv-ce}~=~\num{-.33} is higher than \ac{pmv}~=~\num{-0.23} for \qty{0.3}{\m\per\s}~$<$~\ac{vr}~$\leq$~\qty{0.46}{\m\per\s}.
Both models have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius}.

@@

To continue using the \ac{pmv} model beyond an absolute value of \num{0.5}, more research is needed to determine whether the inaccuracy of the model is related to either or a combination of the following factors and if those sources of inaccuracy can be reduced:
\begin{enumerate}
    \item verify the assumption that the heat losses/gains estimated by the \ac{pmv} are correlated to the thermoregulatory response that the body has to undertake to compensate for unsatisfactory thermal comfort conditions;
    \item if the previous assumption holds, research is needed to validate the correlation coefficients used by Fanger to map the heat losses/gains and thermal stress;
    \item currently a single correlation coefficient is used for the hot and cold side but this may not be a valid assumption and two different sets of equations may be used;
    \item the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
    The \ac{pmv} model assumes that a score of 3 (`hot') signifies the change between a compensable set of conditions to an un-compensable one, e.g., skin wettedness reaches its max value ($w_{max}$), as suggested by Gagge~\cite{GaggeSET}.
    On the contrary, participants indoors may deem the environment to be `hot' as soon as their \ac{w} crosses a critical threshold, which is only a fraction of $w_{max}$.
\end{enumerate}

A combination of all the above-mentioned factors likely plays a role in explaining the discrepancies between the \ac{pmv} model and the self-reported \ac{tsv}.
The latter hypothesis -- 4) anchoring -- has gained significant momentum in the literature, which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equations~\cite{Yao2022, Toftum2002}.
Supporting evidence was provided by \mycite{schweiker2020evaluating}, who showed that the distances between the anchors are not constant and a layperson may not perceive the \ac{tsv} scale as researchers believe they do~\cite{schweiker2019scales, schweiker2020evaluating}.

\paragraph{Measurement Errors}
All the data contained in the \ac{db2} have been published in peer-reviewed manuscripts; however, measurement errors are inevitably present in the dataset.
Instrumentation is not always calibrated and the uncertainty of the measurements may be large, the placement of the instrumentation may not be in the participant's proximity, and estimating clothing and metabolic rates from tables is a complex and prone-to-error task.
Even when variables are measured within the standards' accuracy requirements, the output of the \ac{pmv} value may vary enough to wrongly estimate the correct \ac{tsv}~\cite{dambrosio_alfano_role_2011}.
In an ideal scenario, errors would be randomly distributed and cancel out, not significantly affecting the overall bias of the model but only affecting the standard deviation of the error~\cite{Humphreys2002}.
However, this may not always be the case, for example, if a researcher often underestimates participants' clothing insulation or activity levels.
Moreover, most of the data in the \ac{db2} also only contains entries where the air speed was measured at a single height.
This is a problem since the \ac{pmv} model assumes that the air speed is uniform across the human body, and this is not always the case.
Measuring the air speed at a single height may lead to record higher values than the averaged velocities that are required for the standards’ comfort models.
This is because the anemometer is almost always placed in relatively open space between mid-body and head level, avoiding measuring the usually sheltered lower body level, consequently the associated sensation prediction will be cooler than it should be.
To account for this error, we only included the studies where the air speed was measured at three different heights in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
However, these results were not significantly different from those presented in this paper, and the \ac{pmv-ce} model still had a higher bias than the \ac{pmv} model.

\paragraph{Thermal Sensation Scale}
One minor, but not negligible, source of error is using a discrete scale to assess thermal sensation (most of the \ac{tsv} in the \ac{db2} were collected using a discrete scale) while the \ac{pmv} output is continuous.
For example, even if the model is \qty{100}{\percent} accurate and the estimated \ac{pmv} value is \num{2.5} the person can only report being `warm (\num{2})' or `hot (\num{3})'.
Using a continuous scale, may in principle be a more robust approach since the distances between the anchors are not constant~\cite{schweiker2019scales, schweiker2020evaluating}.
They also showed that the \ac{tsv} assumption of its independence from contextual factors such as climate, season and language is flawed~\cite{schweiker2019scales, schweiker2020evaluating}.
An alternative solution would be to transform the \ac{pmv} output to an ordinal seven-point scale.

>>>>

Figure~\ref{fig:dist_other_data} shows the distribution of age, height, weight, and running mean outdoor temperature.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/dist_other_data}
    \caption{Distribution of age, height, weight, and running mean outdoor temperature.}
    \label{fig:dist_other_data}
\end{figure*}
Less than half (\var{entries_with_sex_info}) of the total entries have information about the participant's sex and are almost equally distributed among males (\qty{52}{\percent}) and females.
Ages are not normally distributed, the same is true for the running mean outdoor temperature.
About half of the participants (\var{percentage_age_20_35}) were aged between \num{20} and \num{35} years old, and only \var{percentage_age_over_60} were older than 60.
This limits our analysis to young adults and does not allow us to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} models in predicting thermal sensation for children or older adults.
The \ac{db2} does not contain information about the health status of the participants.
Approximately \var{running_mean_10_25} of the running mean outdoor temperature values are between \qtyrange{10}{25}{\celsius}.

@@

    Each bin in the right figure has more data points than in the left figure since less data on \ac{tpv} were available.}

@@

Fanger, in his seminal book on thermal comfort, wrote that ``The dissatisfied are defined here as those who vote \num{-2} (cool) or \num{-3} (cold), \num{2} (warm) or \num{3} (hot).

@@

\subsection{Model Bias}\label{subsec:model-bias}
The intended aim of the \ac{pmv} model is not to accurately predict individual thermal responses from participants.
The \ac{pmv} model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.
The results presented so far are primarily comparing individual \ac{tsv} to \ac{pmv}.
This type of analysis does not conclusively disprove the inefficacy of the \ac{pmv} models in predicting the average thermal sensation of a large group of occupants.
Moreover, grouping the results into discrete categories introduces rounding errors since some participants reported \ac{tsv} on a continuous scale and the \ac{pmv} value is a continuous variable.
To compensate for this, we plotted the \ac{pmv} and \ac{pmv-ce} values as a function of \ac{tsv} in Figure~\ref{fig:bubble_models_vs_tsv} and we plotted a \ac{lowess} curve.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/bubble_models_vs_tsv}
    \caption{The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data.
    Raw data were then binned and a bubble chart (circle area is proportional to the number of votes in that bin) was superimposed over the regression curve to aid the visualization of a large dataset.
    The brown dashed line represents the identity line where the slope is 1 and the intercept is 0.}
    \label{fig:bubble_models_vs_tsv}
\end{figure*}
The curve, in the Figure, is calculated using the individual data and not the binned data.
We only binned the data to aid the visualization of a large dataset.
For the model to accurately predict people’s thermal sensation, the regression line (black continuous line) should pass through the origin of the Cartesian plane and have a slope of 1.
This ideal prediction is referred to as the identity line, shown as the brown dashed line in the figure.
The intercepts for the \ac{pmv} is \var{intercept_all_pmv} and for the \ac{pmv-ce} is \var{intercept_all_pmv_ce} which means that the model has a bias of less than one-quarter of a thermal sensation interval.
% todo check data manually
This result is aligned with those presented in Figure~\ref{fig:bar_stacked_model_accuracy} where the \ac{pmv} and \ac{pmv-ce} models can predict people who reported to be neutral \qty{51}{\percent} and \qty{58}{\percent} of the times, respectively.
Both \ac{pmv} formulations always under-predict the thermal sensation of people who reported to be on the extremes of the \ac{tsv};
the slope of the curve is significantly lower than 1.

\subsubsection{Model Overall Bias}\label{subsubsec:model-overall-bias}
We calculated the overall bias of the model as previously done by \mycite{Humphreys2002} by subtracting the \ac{pmv} and \ac{pmv-ce} predictions from the self-reported \ac{tsv}.
Figure~\ref{fig:hist_discrepancies} contains the summary statistics.
% todo check data manually
Overall, both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/hist_discrepancies}
    \caption{Overall bias of the \ac{pmv} and \ac{pmv-ce} models for the subset of data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
    The x-axis shows the discrepancy (bias) between the predicted and self-reported \ac{tsv}.
    We have binned the data into intervals of equal width.
    We colored the bars in the range from \numrange{-.5}{.5} to highlight the number of responses that are within this range.
    We report the median and inter-quantile ranges since data are not normally distributed.
    }
    \label{fig:hist_discrepancies}
\end{figure*}
When including only the data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} the median bias of the \ac{pmv} model reduces to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improves.
The median bias for the \ac{pmv-ce} also decreases to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0} but it is higher than the one for the \ac{pmv}.

\subsubsection{Model Bias as Function of Five Input Variables}\label{subsubsec:model-bias-variable}
Here we present the model bias as a function of \ac{tdb}, \ac{vr}, \ac{rh}, and \ac{met}.
Results are shown in Figure~\ref{fig:bias_models}.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/bias_models}
    \caption{Bias of the \ac{pmv} and \ac{pmv-ce} models plotted as a function of the input variables.
    The x ticks labels show the bin range.
    The square bracket($]$) means that the right value is included in the range.
    The red dashed lines show the acceptable range of bias.
    Each bin has a different number of data points.
    }
    \label{fig:bias_models}
\end{figure*}
In each sub-figure, we have binned the results into discrete intervals.
The x-axis tick label shows the bin range.
All the boxen plots have the same area to aid visualization, even though the number of points in each bin is not constant.
The results depict that the use of the \ac{pmv-ce} model does not improve the bias of the \ac{pmv} results across all input variables.
% todo check data manually
On the contrary, the bias of the \ac{pmv-ce}~=~\num{-.33} is higher than \ac{pmv}~=~\num{-0.23} for \qty{0.3}{\m\per\s}~$<$~\ac{vr}~$\leq$~\qty{0.46}{\m\per\s}.
Both models have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius}.

@@

To continue using the \ac{pmv} model beyond an absolute value of \num{0.5}, more research is needed to determine whether the inaccuracy of the model is related to either or a combination of the following factors and if those sources of inaccuracy can be reduced:
\begin{enumerate}
    \item verify the assumption that the heat losses/gains estimated by the \ac{pmv} are correlated to the thermoregulatory response that the body has to undertake to compensate for unsatisfactory thermal comfort conditions;
    \item if the previous assumption holds, research is needed to validate the correlation coefficients used by Fanger to map the heat losses/gains and thermal stress;
    \item currently a single correlation coefficient is used for the hot and cold side, but this may not be a valid assumption and two different sets of equations may be used;
    \item the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
    The \ac{pmv} model assumes that a score of 3 (`hot') signifies the change between a compensable set of conditions to an un-compensable one, e.g., skin wettedness reaches its max value ($w_{max}$), as suggested by Gagge~\cite{GaggeSET}.
    On the contrary, participants indoors may deem the environment to be `hot' as soon as their \ac{w} crosses a critical threshold, which is only a fraction of $w_{max}$.
\end{enumerate}

A combination of all the above-mentioned factors likely plays a role in explaining the discrepancies between the \ac{pmv} model and the self-reported \ac{tsv}.
The latter hypothesis -- 4) anchoring -- has gained significant momentum in the literature, which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equations~\cite{Yao2022, Toftum2002}.
Supporting evidence was provided by \mycite{schweiker2020evaluating}, who showed that the distances between the anchors are not constant, and a layperson may not perceive the \ac{tsv} scale as researchers believe they do~\cite{schweiker2019scales, schweiker2020evaluating}.

\paragraph{Measurement Errors}
All the data contained in the \ac{db2} have been published in peer-reviewed manuscripts; however, measurement errors are inevitably present in the dataset.
Instrumentation is not always calibrated and the uncertainty of the measurements may be large, the placement of the instrumentation may not be in the participant's proximity, and estimating clothing and metabolic rates from tables is a complex and prone-to-error task.
Even when variables are measured within the standards' accuracy requirements, the output of the \ac{pmv} value may vary enough to wrongly estimate the correct \ac{tsv}~\cite{dambrosio_alfano_role_2011}.
In an ideal scenario, errors would be randomly distributed and cancel out, not significantly affecting the overall bias of the model but only affecting the standard deviation of the error~\cite{Humphreys2002}.
However, this may not always be the case, for example, if a researcher often underestimates participants' clothing insulation or activity levels.
Moreover, most of the data in the \ac{db2} also only contains entries where the air speed was measured at a single height.
This is a problem since the \ac{pmv} model assumes that the air speed is uniform across the human body, and this is not always the case.
Measuring the air speed at a single height may lead to recording higher values than the averaged velocities that are required for the standards’ comfort models.
This is because the anemometer is almost always placed in a relatively open space between mid-body and head level, avoiding measuring the usually sheltered lower body level, consequently, the associated sensation prediction will be cooler than it should be.
To account for this error, we only included the studies where the air speed was measured at three different heights in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
However, these results were not significantly different from those presented in this paper, and the \ac{pmv-ce} model still had a higher bias than the \ac{pmv} model.

\paragraph{Thermal Sensation Scale}
One minor, but not negligible, source of error is using a discrete scale to assess thermal sensation (most of the \ac{tsv} in the \ac{db2} were collected using a discrete scale) while the \ac{pmv} output is continuous.
For example, even if the model is \qty{100}{\percent} accurate and the estimated \ac{pmv} value is \num{2.5} the person can only report being `warm (\num{2})' or `hot (\num{3})'.
Using a continuous scale, may in principle be a more robust approach since the distances between the anchors are not constant~\cite{schweiker2019scales, schweiker2020evaluating}.
They also showed that the \ac{tsv} assumption of its independence from contextual factors such as climate, season, and language is flawed~\cite{schweiker2019scales, schweiker2020evaluating}.
An alternative solution would be to transform the \ac{pmv} output to an ordinal seven-point scale.
